---
title: Syllabus
slug: syllabus
citeproc: false
bibliography: ../static/bib/references.bib
csl: ../static/bib/chicago-syllabus-no-bib.csl
---

```{r load-libraries, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tibble)
library(pander)
```

- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/files/syllabus.pdf)

> "The consequences for human welfare involved in questions like these are simply staggering: Once one starts to think about them, it is hard to think about anything else. -- Robert Lucas, Nobel Laureate in Economics 1995"

<span class="newthought">Economic Development</span> [^courseinfo] asks and attempts to answer the fundamental question posed by Adam Smith in 1776: *why are some countries wealthy, and others poor?*. Before "economics" emerged as its own discipline in the 20^th^ century, and adopted more rigorous mathematical trappings, the exploration of *political economy* attempted to answer this question: political, social, and cultural institutions and histories had as much to offer as the division of labor and market exchanges in answering the challenge. In the aftermath of WWII, economists in "the First World" began more consciously studying how to promote development in "the Third World", while policymakers built international institutions (the IMF, the World Bank, the Bretton-Woods financial systems) aimed at securing peace and outwardly developing what they say as the "Third World." In order to grapple with these key questions, we will examine a mixture of political economy and economic history to understand the role of political, cultural, and social institutions in directing economic activity towards prosperity or towards ruin. This course, like the professionals dealing with the big questions, will offer many suggestions but fewer "correct" or concrete answers than you may be used to. You should come to this course as a willing participant in the ongoing conversation.

The economics of development combines core themes and models of macroeconomics (growth theory, macroeconomic stability and policy) with core principles of microeconomics (price theory), and as such, the **prerequisites** for this course are **either ECON 205 or ECON 206**.

This course will be a hybrid of formal lecture, hands-on activities, and student presentations of scholarly articles, with exams and a writing writing assignment serving as the prime means of evaluation. 

[^courseinfo]: {-}
  `r blogdown::shortcode("courseinfo")`


# Student Learning objectives

<span class="newthought">By the end of this course,</span> you will: 

- Explain how the development community measures economic development
- Interpret regression tables in the empirical literature in development
- Demonstrate different theories of economic development
- Explain why various policies aimed at promoting development have failed
- Describe essential conditions for successful development
- Discuss the broad economic history of "the West", several key "Emerging Markets" (such as Russia, China, Korea, etc.), and several other case studies of developing countries

Given these objectives, this course fulfills two of the learning outcomes for [the George B. Delaplaine, Jr. School of Business](https://www.hood.edu/academics/departments/george-b-delaplaine-jr-school-business/student-learning-outcomes) Economics B.A. program:

- Apply economic reasoning and models to understand and analyze problems of public policy [...]
- Demonstrate effective oral and written communications skills for personal and professional success[...]

<span class="newthought">My standard disclaimer:</span> This class may challenge many of your existing beliefs and conceptions about how the world works, and how it should work. This is the most important and exciting part of a liberal arts education. This does *not* mean that I want to make you to see everything "my way." In fact, if you come out of this class thinking exactly like me, then I have probably failed you as a teacher. To the best of my ability, I keep my opinions to myself unless relevant for purposes of discussion, and I respect and invite you to reach your own conclusions on all matters. 

<span class="newthought">Content warning:</span> this class will cover sensitive political and cultural topics and compel you to grapple with countries, cultures, and viewpoints very different from your own. To put it mildly, these topics may include themes of violence, slavery, imperialism, and different ideologies inherently wrapped up in the tragic history of both the developed and developing world.

If at any point you find yourself struggling in this course for any reason, please come see me. Do not suffer in silence! Coming to see me for help does not diminish my view of you, in fact I will hold you in *higher* regard for understanding your own needs and taking charge of your own learning. There are also a some fantastic resources on campus, such as the [Center for Academic Achievement and Retention (CAAR)](http://www.hood.edu/campus-services/academic-services/index.html) and the [Beneficial-Hodson Library](http://www.hood.edu/library/). 

# Course materials

<span class="newthought">Most of the readings in this class are free.</span> 

## Books

The following book is **required**^[You are not *obligated* to buy it, I just **strongly recommend** it in the sense that you will still have access to all data and assignments without possessing the book. But this is a course where you really will want to understand the derivations or get additional context beyond just my slides...] and will be available from the campus bookstore.

- [@Bailey], 2^nd^ ed. ISBN: 978-0190857462


## Articles

Throughout the course, I will post both required and supplemental (non-required) readings that enrich your understanding for each topic on Blackboard. Check Blackboard *frequently* for announcements and updates to assignments, readings, and grades.


# Assignments and grades

<span class="newthought">You can find descriptions</span> for all the assignments on the [assignments page](/assignments/).

```{r assignments-grades, include=FALSE}
assignments <- tribble(
  ~"",  ~Assignment, ~Percent,
  "n"     , "Article Summary (Average)", "5%",
  1, "Country Profile", "5%",
  2      , "Short Papers", "10% each",
  1     , "Midterm", "25%",
  1     , "Final", "25%"
)

grading <- tribble(
  ~Grade, ~Range, ~Grade1, ~Range1,
  "A",  "93–100%", "C",  "73–76%",
  "A−", "90–92%", "C−", "70–72%",
  "B+", "87–89%", "D+", "67–69%",
  "B",  "83–86%", "D",  "63–66%",
  "B−", "80–82%", "D−", "60–62%",
  "C+", "77–79%", "F",  "< 60%"
)
```

```{r}
knitr::kable(assignments)
```

```{r show-assignments-table, eval=F, echo=FALSE, results="asis"}
assignments_total <- tibble(Assignment = "Total",
                            Points = sum(assignments$Points),
                            Percent = NA)

bind_rows(assignments, assignments_total) %>% 
  select(Assignment, Points, Percent) %>% 
  pandoc.table(justify = "lcc", missing = "—")
```

\

```{r show-grades-table, echo=FALSE, results="asis"}
grading %>%
  pandoc.table(justify = "lclc", 
               col.names = c("Grade", "Range", "Grade", "Range"))
```

# Policies and Expectations

### Attendance and Participation

I expect you to attend class and to come having already done the reading assigned for that day. I will remind you in class and possibly through Blackboard or email which readings I want you to read for the next class. You are all adults and I will treat you as such. I do not take attendance, nor do I grade formally for participation but I strongly recommended you attend class and participate for your sake and the sake of your classmates. If you are too distracted or are not prepared to learn, I suggest you stay home, where you can check Facebook more efficiently. I reserve the right to boost the final grades of students that I believe have made consistent, quality contributions above and beyond their peers in class conversations by up to 2.5 points. 

### Absences and Make-Ups

You generally do *not* need to let me know if you are unable to make class, *unless* it is on the day of an exam. It will however, be your responsibility to acquire the notes from a classmate for any missed classes. If you are unable to attend an exam for a legitimate reason (e.g. sports/club events, traveling, illness, family issues), please notify me at least *one week* in advance, and we will schedule a make-up exam date. If you are ill or otherwise unable to attend on the day of the exam, contact me ASAP to make arrangements. Failure to do so, including desperate attempts to make arrangements only *after* the absence will result in a grade of 0 and little sympathy. I reserve the right to re-weight other assignments for students who I believe are legitimately unable to complete a particular assignment.

### Email Accounts

Students must regularly monitor their Hood email accounts to receive important college information, including messages related to this class. Email through the Blackboard system is my main method of communicating announcements and deadlines regarding your assignments. **Do not reply to the automated Blackboard email**. My Hood email (safner@hood.edu) is the best means of contacting me. I will do my best to respond within 24 hours. If I do not reply within 48 hours, do not take it personally, and *feel free to send a follow up email* in case I genuinely missed your original message. 

### Office Hours

I am generally in my office Monday-Thursday during "normal business hours." You are always welcome to walk-in and chat about class, college, careers, or whatever you like. Please try to use the official office hours stated at the head of the syllabus if possible. If you need to meet at a different time, I request that you send me an email or let me know after class so I know when to expect you. If you want to go over material from class, please have *specific* questions you want answered. 

### Enrollment

Students are responsible for verifying their enrollment in this class.  The last day to add or drop this class \textbf{Tuesday, September 4}.

### Honor Code

Hood College has an Academic Honor Code which requires all members of this community to maintain the highest standards of academic honesty and integrity. Cheating, plagiarism, lying, and stealing are all prohibited.  All violations of the Honor Code are taken seriously, will be reported to appropriate authority, and may result in severe penalties, including expulsion from the college. See [here](http://hood.smartcatalogiq.com/en/2016-2017/Catalog/The-Spirit-of-Hood/The-Academic-Honor-Code-and-Code-of-Conduct) for more detailed information.

### Accessibility and Students With Disabilities

This course is intended to be accessible for all students, including those with mental, physical, or cognitive disabilities, illness, injuries, impairments, or any other condition that tends to negatively affect one's equal access to education. If at any point in the term, you find yourself not able to fully access the space, content, and experience of this course, you are welcome (and not required) to contact me by email, phone, or during office hours to discuss your specific needs. I also encourage you to contact the Office of Accessibility Services (301-696-3421 or gmuer@hood.edu). If you have a diagnosis or history of accommodations in high school or previous postsecondary institutions, Accessibility Services can help you document your needs and create an accommodation plan. By making a plan through Accessibility Services, you can ensure appropriate accommodations without disclosing your condition or diagnosis to course instructors.  


# Schedule

<span class="newthought">You can find a full roadmap</span> with resources for each class meeting on the [schedule page](/schedule/).
